---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
fatal_police_shootings <- read_csv("data/fatal-police-shootings-data.csv")

# Clean required data and prepare for analysis if needed. 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [“Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. 

But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot 

and killed and 28 percent of the unarmed deaths”]

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
black_females_shot <- fatal_police_shootings %>% filter(gender == "F" & race == "B")
#number of total rows= 7139
#number of black females shot= 54
54/7139 *100 
#Percentage of Black women shot was 0.76%
```
```{r}
women_shot <- fatal_police_shootings %>% filter(gender == "F")
  #number of women = 320
54/320*100
#percentage was 16.88%
```
```{r}
unarmed_shootings <- fatal_police_shootings %>% filter(armed == "unarmed")
#number is 442
blackfem_unarmed_shootings <- unarmed_shootings %>% filter(gender == "F" & race == "B")
7/442*100
###come back
```
-30-