---
title: "lab_03"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---   

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)
# Load the tidyverse here
library(tidyverse)
# Load janitor here
install.packages('janitor')
```

## Load data
```{r}
# Write code to load the WV slice of the ppp loan data here and call it wv_ppp_loans.
# The data is called lab_03.rds, and is in the same folder as lab_03.Rmd (this data notebook)

wv_ppp_loans <- read_rds("lab_03.rds")

```

## Answer questions

Q1: Which five cities located in Kanawha County, West Virginia, got the largest _number_ of loans? How many did each get? To answer this question, you'll have to clean the very messy "city" field.

Note: To answer this question, you'll need to see every variation of the "city" column in Kanawha County, and then fully clean the city field for those cities. Hint: once cleaned, the dataframe that counts the number of loans by city should have 38 distinct cities for Kanawha County.

A1:Charleston(2802), South Charleston(478), St. Albans(296), Dunbar(162) and Nitro (140) had the largest number of loans respectively. 
```{r}
kanawha_county <- wv_ppp_loans %>% filter(project_county_name == "KANAWHA") %>% 
mutate(city = str_to_title(city)) %>%
mutate(city = case_when(
    city == "Balle" ~ "Belle",
    TRUE ~ city
  ))  %>% 
mutate(city = case_when(
    city == "Charelston" ~ "Charleston",
    TRUE ~ city
  ))  %>% 
mutate(city = case_when(
    city == "Crosslanes" ~ "Cross Lanes",
    TRUE ~ city
  ))  %>% 
mutate(city = case_when(
    city == "Elkveiw" ~ "Elkview",
    TRUE ~ city
  ))  %>% 
mutate(city = case_when(
    city == "S Charleston" ~ "South Charleston",
    TRUE ~ city
  ))  %>% 
mutate(city = case_when(
    city == "So Charleston" ~ "South Charleston",
    TRUE ~ city
  ))  %>% 
mutate(city = case_when(
    city == "Saint Albans" ~ "St. Albans",
    TRUE ~ city
  ))  %>% 
mutate(city = case_when(
    city == "St Albans" ~ "St. Albans",
    TRUE ~ city
  ))  %>% 
mutate(city = case_when(
    city == "St.albans" ~ "St. Albans",
    TRUE ~ city
  ))  %>% 
group_by(city) %>% 
select(city, amount) %>% 
summarise(total=n())

```

Q2: Find the 5-digit ZIP code in West Virginia that had the highest total amount of loans (by dollar value). What is the name of the business that got the largest loan (by dollar value) in that zip code? Write up a short paragraph that also answers these questions:

* Where in West Virginia is this ZIP code located?
* How much was the business' loan for?
* How much of that loan was forgiven?
* Which number is bigger, the loan amount or the forgiveness amount? Can you explain the discrepancy?

A2:The zip code with the highest amount of loans was 25301, which is Kanahaw County in western WV. The business in the zip code that had the highest loan THE EYE AND EAR CLINIC OF CHARLESTON, which took a loan of 1,520,007.00. The loan forgiveness amount was 18,100.73, making the loan much bigger than the forgiveness amount. This could be due to the company not following the SBA's guidelines for loan forgiveness.

```{r}
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount
# Second, use the top ZIP code to create a separate dataframe that shows the business with the biggest loan in that ZIP code.
wv_ppp_loans %>% 
mutate(zip = str_sub(zip, start=1L, end=5L)) %>% 
group_by(zip) %>% 
summarise(total=sum(amount)) %>% 
arrange(desc(total))
```
```{r}
largest_zip <- wv_ppp_loans %>% filter(zip == "25301") 

```
Q3: Five business entities owned by the spouse of a member of West Virginia's U.S. House delegation got PPP loans. Write up a paragraph that answers the following questions.   

  * What are the businesses' names?
  * What do the businesses do?
  * How much money did the five businesses receive in total?
  * What percentage of the total loan amount have been forgiven?
  * How did the U.S. House member vote on the legislation initially authorzing the PPP loan program.

Feel free to add any interesting "color" (little interesting details) you find in your research.  

A3:The five businesses are DM Motors Inc., Dutch Miller Chevrolet, Dutch Miller of Charleston, Dutch Miller of Charlotte and Dutch Miller Subaru. DM Motors Inc. is the company that manages these other companies, which are car dealerships. The companies received $3,105,300 total in loans. 101.13% of the total loan was forgiven. Rep. Carol Miller voted to pass the program.


```{r}
# Hint
# First, create a dataframe that only has loans to those five companies
# Second, use that newly created dataframe to calculate total loan amount and percentage forgiven for those companies

dutch_miller <- wv_ppp_loans %>% 
filter (str_detect(name, "DUTCH MILLER |DM MOTOR")) %>% 
select(name, amount, forgiveness_amount) %>%
mutate( percent_forgiven = (forgiveness_amount/amount)*100) %>% 
arrange(percent_forgiven, desc(amount))

```
